---
title: "Diamonds"
author: "Samuel"
date: "2016Äê12ÔÂ22ÈÕ"
output: html_document
---

```{r download diamonds}
library(ggplot2)
data(diamonds)

summary(diamonds)
```
# Create a histogram of the price of
# all the diamonds in the diamond data set.

```{r}
qplot(x = price, data = diamonds)


sum(ifelse(diamonds$price < 500,1,0))

sum(ifelse(diamonds$price < 250,1,0))

sum(ifelse(diamonds$price >= 15000,1,0))
```

# Explore the largest peak in the
# price histogram you created earlier.

# Try limiting the x-axis, altering the bin width,
# and setting different breaks on the x-axis.

# There won¡¯t be a solution video for this
# question so go to the discussions to
# share your thoughts and discover
# what other people find.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).


```{r}
qplot(x = price, data = diamonds, binwidth = 20) +
  scale_x_continuous(limits = c(300, 1500),
                     breaks = seq(300,1500,100))
```

# Break out the histogram of diamond prices by cut.

# You should have five histograms in separate
# panels on your resulting plot.
```{r}
library(gridExtra) 

qplot(x = price, data = diamonds) +
  facet_wrap(~diamonds$cut, scales = "free_y", ncol = 3)

```

```{r}
by(diamonds$price, diamonds$cut, summary)
```

# Create a histogram of price per carat
# and facet it by cut. You can make adjustments
# to the code from the previous exercise to get
# started.

# Adjust the bin width and transform the scale
# of the x-axis using log10.

```{r}
qplot(x = log10(price/carat +1), data = diamonds, binwidth = 0.1) +
  facet_wrap(~diamonds$cut, scales = "free_y", ncol = 5)

```

# Investigate the price of diamonds using box plots,
# numerical summaries, and one of the following categorical
# variables: cut, clarity, or color.

# There won¡¯t be a solution video for this
# exercise so go to the discussion thread for either
# BOXPLOTS BY CLARITY, BOXPLOT BY COLOR, or BOXPLOTS BY CUT
# to share you thoughts and to
# see what other people found.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).
```{r}
p1 <- qplot(x = cut, y = price,
          data = subset(diamonds, !is.na(cut)), 
          geom = 'boxplot') +
        scale_y_continuous(limits = c(0, 6000))

p2 <- qplot(x = clarity, y = price,
          data = subset(diamonds, !is.na(cut)), 
          geom = 'boxplot') +
        scale_y_continuous(limits = c(0, 6000))

p3 <- qplot(x = color, y = price,
          data = subset(diamonds, !is.na(cut)), 
          geom = 'boxplot') +
        scale_y_continuous(limits = c(0, 6000))

grid.arrange(p1, p2, p3, ncol = 1)

by(diamonds$price, diamonds$color, summary)

```
# Investigate the price per carat of diamonds across
# the different colors of diamonds using boxplots.
```{r}
qplot(x = color, y = price/carat, data = diamonds, geom = 'boxplot') 

```


```{r Frequency Polygons}
qplot(x = carat, 
        data = diamonds,
        xlab = 'The weight of diamonds',
        ylab = 'Proportion of diamonds with that weight',
        geom = 'freqpoly',
        binwidth = 0.1) +
  scale_x_continuous() 
```

# Your first task is to create a
# scatterplot of price vs x.
# using the ggplot syntax.

```{r}
ggplot(aes(x = x, y = price), data = diamonds) + geom_point()

```

# The coorelation between price and x,y,z
```{r}
cor.test(diamonds$price, diamonds$x)

with(diamonds, cor.test(price, y))

with(diamonds, cor.test(price, z))
```

# Create a simple scatter plot of price vs depth.
```{r}
ggplot(aes(x = depth, y = price), data = diamonds) + 
  geom_point(alpha = 0.01) + 
  scale_x_continuous(breaks = seq(50, 70, 2), lim = c(50, 70))
```

# Create a scatterplot of price vs carat
# and omit the top 1% of price and carat
# values.
```{r}
ggplot(aes(x = carat, y = price), data = diamonds) + 
  geom_point(alpha = 0.01) + 
  xlim(0, quantile(diamonds$carat, 0.99)) +
  ylim(0, quantile(diamonds$price, 0.99)) 
```

# Create a scatterplot of price vs. volume (x * y * z).
# This is a very rough approximation for a diamond's volume.

# Create a new variable for volume in the diamonds data frame.
# This will be useful in a later exercise.
```{r}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z

ggplot(aes(x = volume, y = price), data = diamonds) + 
  geom_point(alpha = 0.01) 
```
```{r}
with(subset(diamonds,volume > 0 & volume < 800), cor.test(price, volume))
```

# Subset the data to exclude diamonds with a volume
# greater than or equal to 800. Also, exclude diamonds
# with a volume of 0. Adjust the transparency of the
# points and add a linear model to the plot. (See the
# Instructor Notes or look up the documentation of
# geom_smooth() for more details about smoothers.)

# We encourage you to think about this next question and
# to post your thoughts in the discussion section.

# Do you think this would be a useful model to estimate
# the price of diamonds? Why or why not?

```{r}
ggplot(aes(x = volume, y = price), data = subset(diamonds, volume > 0 & volume < 800)) + 
  geom_point(alpha = 0.01) +
  geom_smooth(method = 'lm', color = 'red') 
```

```{r}
diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarise(mean_price = mean(price),
            median_price = median(price),
            min_price = min(price),
            max_price = max(price),
            n = n()) %>%
  arrange(clarity)
```

```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

p1 <- qplot(data = diamonds_mp_by_clarity, mean_price) + geom_bar()
p2 <- qplot(data = diamonds_mp_by_color, mean_price) + geom_bar()
grid.arrange(p1, p2, ncol = 1)
```
