---
title: "Diamonds"
author: "Samuel"
date: "2016Äê12ÔÂ22ÈÕ"
output: html_document
---

```{r download diamonds}
library(ggplot2)
data(diamonds)

summary(diamonds)
```
# Create a histogram of the price of
# all the diamonds in the diamond data set.

```{r}
qplot(x = price, data = diamonds)


sum(ifelse(diamonds$price < 500,1,0))

sum(ifelse(diamonds$price < 250,1,0))

sum(ifelse(diamonds$price >= 15000,1,0))
```

# Explore the largest peak in the
# price histogram you created earlier.

# Try limiting the x-axis, altering the bin width,
# and setting different breaks on the x-axis.

# There won¡¯t be a solution video for this
# question so go to the discussions to
# share your thoughts and discover
# what other people find.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).


```{r}
qplot(x = price, data = diamonds
      )
```












