---
title: "Diamonds"
author: "Samuel"
date: "2016Äê12ÔÂ22ÈÕ"
output: html_document
---

```{r download diamonds}
library(ggplot2)
data(diamonds)

summary(diamonds)
```
# Create a histogram of the price of
# all the diamonds in the diamond data set.

```{r}
qplot(x = price, data = diamonds)


sum(ifelse(diamonds$price < 500,1,0))

sum(ifelse(diamonds$price < 250,1,0))

sum(ifelse(diamonds$price >= 15000,1,0))
```

# Explore the largest peak in the
# price histogram you created earlier.

# Try limiting the x-axis, altering the bin width,
# and setting different breaks on the x-axis.

# There won¡¯t be a solution video for this
# question so go to the discussions to
# share your thoughts and discover
# what other people find.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).


```{r}
qplot(x = price, data = diamonds, binwidth = 20) +
  scale_x_continuous(limits = c(300, 1500),
                     breaks = seq(300,1500,100))
```

# Break out the histogram of diamond prices by cut.

# You should have five histograms in separate
# panels on your resulting plot.
```{r}
library(gridExtra) 

qplot(x = price, data = diamonds) +
  facet_wrap(~diamonds$cut, scales = "free_y", ncol = 3)

```

```{r}
by(diamonds$price, diamonds$cut, summary)
```

# Create a histogram of price per carat
# and facet it by cut. You can make adjustments
# to the code from the previous exercise to get
# started.

# Adjust the bin width and transform the scale
# of the x-axis using log10.

```{r}
qplot(x = log10(price/carat +1), data = diamonds, binwidth = 0.1) +
  facet_wrap(~diamonds$cut, scales = "free_y", ncol = 5)

```

# Investigate the price of diamonds using box plots,
# numerical summaries, and one of the following categorical
# variables: cut, clarity, or color.

# There won¡¯t be a solution video for this
# exercise so go to the discussion thread for either
# BOXPLOTS BY CLARITY, BOXPLOT BY COLOR, or BOXPLOTS BY CUT
# to share you thoughts and to
# see what other people found.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).
```{r}
p1 <- qplot(x = cut, y = price,
          data = subset(diamonds, !is.na(cut)), 
          geom = 'boxplot') +
        scale_y_continuous(limits = c(0, 6000))

p2 <- qplot(x = clarity, y = price,
          data = subset(diamonds, !is.na(cut)), 
          geom = 'boxplot') +
        scale_y_continuous(limits = c(0, 6000))

p3 <- qplot(x = color, y = price,
          data = subset(diamonds, !is.na(cut)), 
          geom = 'boxplot') +
        scale_y_continuous(limits = c(0, 6000))

grid.arrange(p1, p2, p3, ncol = 1)

by(diamonds$price, diamonds$color, summary)

```
# Investigate the price per carat of diamonds across
# the different colors of diamonds using boxplots.
```{r}
qplot(x = color, y = price/carat, data = diamonds, geom = 'boxplot') 

```


```{r Frequency Polygons}
qplot(x = carat, 
        data = diamonds,
        xlab = 'The weight of diamonds',
        ylab = 'Proportion of diamonds with that weight',
        geom = 'freqpoly',
        binwidth = 0.1) +
  scale_x_continuous() 
```


