---
title: "Prosper Loan Exploration"
author: "Samuel Wang"
date: "2017-01-05"
output: html_document
---

What borrower rate you need to pay when you wanna a loan from Prosper
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.
install.packages("ggplot2", dependencies = T) 
install.packages("knitr", dependencies = T)
install.packages("dplyr", dependencies = T)

library(ggplot2)
library(plyr)
library(dplyr)
```
This report explores a dataset containing LoanStatus and attributes for almost 114000 loan data in Prosper. I pick out 15 variables from the total 81 variables which I think are strongly relative to borrower. So firstly I deal with the original data into a subset and take a look at the subset as follows:
```{r echo=FALSE, Load_the_Data}
# Load the Data
prosperLoan <- read.csv(file = "C:\\Users\\wangs\\Downloads\\subsetProsperLoanData.csv", header = TRUE, sep = ",")

subvariables <- c('ListingCreationDate','CreditGrade','Term','LoanStatus','BorrowerRate','ProsperRating..Alpha.','ProsperScore','ListingCategory..numeric.','BorrowerState','Occupation','CurrentDelinquencies','IncomeRange','TotalProsperLoans','TotalProsperPaymentsBilled','OnTimeProsperPayments','LoanOriginalAmount')

prosperLoan <- subset(prosperLoan,select=subvariables)

colnames(prosperLoan)[6] <- "ProsperRating"
colnames(prosperLoan)[8] <- "ListingCategory"

prosperLoan$TotalProsperPaymentdelinquencies <- prosperLoan$TotalProsperPaymentsBilled - prosperLoan$OnTimeProsperPayments

subvariables <- c('ListingCreationDate','CreditGrade','Term','LoanStatus','BorrowerRate','ProsperRating','ProsperScore','ListingCategory','BorrowerState','Occupation','CurrentDelinquencies','IncomeRange','TotalProsperLoans','TotalProsperPaymentdelinquencies','LoanOriginalAmount')

prosperLoan <- subset(prosperLoan,select=subvariables)

prosperLoan[is.na(prosperLoan)] <- 0

ListingCategoryMap <- c("0"="Not Available","1"="Debt Consolidation","2"="Home Improvement","3"="Business",
                        "4"="Personal Loan","5"="Student Use","6"="Auto","7"="Other","8"="Baby&Adoption",
                        "9"="Boat","10"="Cosmetic Procedure","11"="Engagement Ring","12"="Green Loans",
                        "13"="Household Expenses","14"="Large Purchases","15"="Medical/Dental","16"="Motorcycle",
                        "17"="RV","18"="Taxes","19"="Vacation","20"="Wedding Loans")

prosperLoan$ListingCategory <- as.character(prosperLoan$ListingCategory)
revalue(prosperLoan$ListingCategory,ListingCategoryMap)

prosperLoan$ListingCategory[prosperLoan$ListingCategory == "0"] <- "Not Available"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "1"] <- "Debt Consolidation"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "2"] <- "Home Improvement"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "3"] <- "Business"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "4"] <- "Personal Loan"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "5"] <- "Student Use"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "6"] <- "Auto"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "7"] <- "Other"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "8"] <- "Baby&Adoption"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "9"] <- "Boat"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "10"] <- "Cosmetic Procedure"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "11"] <- "Engagement Ring"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "12"] <- "Green Loans"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "13"] <- "Household Expenses"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "14"] <- "Large Purchases"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "15"] <- "Medical/Dental"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "16"] <- "Motorcycle"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "17"] <- "RV"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "18"] <- "Taxes"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "19"] <- "Vacation"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "20"] <- "Wedding Loans"

prosperLoan$ListingCategory <- as.factor(prosperLoan$ListingCategory)

prosperLoan$ListingCreationDate <- as.Date(prosperLoan$ListingCreationDate, format='%Y-%m-%d')

write.csv(file = './subsetProsperLoanData.csv', prosperLoan, na="0")

names(prosperLoan)
```
# Univariate Plots Section
The variable 'TotalProsperPaymentdelinquencies' represents the delinquencies a borrower made, which is equal to the difference between 'TotalProsperPaymentsBilled' and 'OnTimeProsperPayments', because I think this variable may represent a borrower's credit, which is very important for applying a loan. Also I transform the value of 'ListingCategory' into the category which it means. And transform the value of 'ListingCreationDate' into date value.

Now let's explore the data. Because I want to disscuss about the borrowers, so the first thing came up to my mind is how easily one person can borrow a loan from Prosper. Then you need know whether an application for a loan is accepted or not, Then I find a variable called LoanStatus.
We can see that most of the loans are active but there still are some defaluted payments. But this variable stands for the status of a loan, I can't find any evidence of how easily one can borrow a loan, so I choose the variable BorrowerRate, which is also very important for a borrower.
```{r echo=FALSE, Load_the_Data}
p <- ggplot(prosperLoan, aes(LoanStatus))
p <- p + geom_bar(width = 1, colour="white", fill="blue")
p <- p + geom_text(aes(y = ..count.., label = ..count..),
                   stat = "count", color = "white",
                   hjust = 1.0, size = 3)
p <- p + theme(legend.position = "none")
p <- p + coord_flip()
p
```
```{r echo=FALSE, Load_the_Data}
ggplot(data=prosperLoan,aes(BorrowerRate)) + 
  geom_histogram(color=I('black'),fill=I('orange'), bins = 50)
```
The Borrower Rate seems quite uniformly distributed, except for the data around 0.32, it seems there are some outliners. so let's take a close look at the range between 0.3-0.4 with 150 bins.
```{r echo=FALSE, Load_the_Data}
ggplot(data=subset(prosperLoan, between(prosperLoan$BorrowerRate, 0.3, 0.4)),aes(BorrowerRate)) + 
  geom_histogram(color=I('black'),fill=I('orange'), bins = 150) + 
  ggtitle('Frequency of Interest Rate (with 150 bins)')
```
While taking the fact that this data is provided by Prosper into consideration, there must be few wrong data. Maybe there is a period when Prosper added thier fee of providing loans, then I draw a histogram to see the user counts at different years. 
```{r echo=FALSE, Load_the_Data}
qplot(x = ListingCreationDate, data = prosperLoan, bins = 120)
```
We could see that the loan records are distributed from the year 2006 to 2014, while there is a gap at 2009, from [Wiki](https://en.wikipedia.org/wiki/Prosper_Marketplace#2009_post_SEC_relaunch), we konw this 'quite period' is because a class action lawsuit was filed against Prosper in the Superior Court of California, County of San Francisco, California. Also I see that the records are much more in 2014 than the former years. With more and more people know Prosper, it becomes popular to invest and borrow a loan on Prosper.

Maybe there is a relationship between the month and the loan record, so I divide the variable 'ListingCreationDate' into year and month to see what will show. 
```{r echo=FALSE, Load_the_Data}
prosperLoan$year <- as.numeric(format(prosperLoan$ListingCreationDate, '%Y'))
prosperLoan$month <- as.numeric(format(prosperLoan$ListingCreationDate, '%m'))

qplot(x = month, data = prosperLoan) +
  scale_x_continuous(breaks = seq(0,12,3)) +
  facet_wrap(~prosperLoan$year, scales = "free_y", ncol = 5)
```
Well, it seems there is no clear evidence.Now we know some information about the loan creation date, and it may influence the borrowerrate. I need analysis more variables which may influence the borrowerRate, and in fact I have zoomed in the variables which I think will have impact on borrowerRate, what I should do now is analysis one by one. Well, the next is the length of the loan expressed in months. 
```{r echo=FALSE, Load_the_Data}
prosperLoan$Term<-factor(prosperLoan$Term)
ggplot(data=prosperLoan,aes(Term)) + 
  geom_bar(color=I('black'),fill=I('blue'))
```
From the above histogram, we can find that Prosper provides mostly 36-months loans, then 60-months, 12-months at least. From the count of loans and rule of thumb, maybe these 12-months loans is those high borrowerRate loans, which we can discuss in the next part. Let's see how custom risk score distribute, which was created by Prosper in order to assess risk. The score ranges from 1-11, with 11 being the best, or lowest, risk score.
```{r echo=FALSE, Load_the_Data}
ggplot(aes(x = ProsperScore), data = prosperLoan) + 
  geom_histogram() +
  scale_x_continuous(breaks = seq(1,11,1))
```
We can see that except for the missing values, borrowers' score are approximately normally distributed. Most of the borrowers have score ranging from 4 to 8. How about the ListingCategory:
```{r echo=FALSE, Univariate_Plots}
prosperLoan <- within(prosperLoan,
                      ListingCategory <- factor(ListingCategory, 
                                                levels = names(sort(table(ListingCategory),decreasing = TRUE))))
p <- ggplot(prosperLoan, aes(ListingCategory))
p <- p + geom_bar(width = 1, colour="white")
p <- p + geom_text(aes(y = ..count.., label = ..count..),
                   stat = "count", color = "white",
                   hjust = 1.0, size = 3)
p <- p + theme(legend.position = "none")
p <- p + coord_flip()
p
```
From the above figure, we could tell that most loans on Prosper are for Debt Consolidation. Excluding the Not Available and Other purposes, most of the loans go into Home Improvement, Business. While categories like entaimment things such as RV, Boat, Motorcycle, Vacation only accupate a small portion of the loan. Therefore, I conclude that most the people borrow money for necessary needs. Then how about the state:
```{r echo=FALSE, Univariate_Plots}
prosperLoan <- within(prosperLoan,
                      BorrowerState <- factor(BorrowerState, 
                                                levels = names(sort(table(BorrowerState),decreasing = TRUE))))
par(mar = c(5, 1, 1, 5), mgp=c(3,1,0),las=0)

ggplot(prosperLoan, aes(BorrowerState)) + 
  geom_bar(bin = 1, colour="white") +
  geom_text(aes(y = ..count.., label = ..count..),
                   stat = "count", color = "white",
                   hjust = 1.0, size = 3) + 
  theme(axis.text.x  = element_text(angle=90, vjust=0.5, size = 7)) 


```
As we can see from this histgrom, CA, TX, NY, Fl and IL have the most borrowers. It seems that these states are the most populous and economically areas as well. How about the occupation of the borrowers:
```{r echo=FALSE, Univariate_Plots}
prosperLoan <- within(prosperLoan,
                      Occupation <- factor(Occupation, 
                                                levels = names(sort(table(Occupation),decreasing = TRUE))))
ggplot(prosperLoan, aes(Occupation)) + 
  geom_bar(width = 1, colour="white") + 
  geom_text(aes(y = ..count.., label = ..count..),
                   stat = "count", color = "white",
                   hjust = 1.0, size = 3) + 
  theme(axis.text.x  = element_text(angle=90, vjust=0.5, size = 7)) 
```
There are so many borrowers who hadn't provided their occupation, ignoring this, Professional, computer programmer, Executive etc. borrow a loan most, and the tail show that few student borrow a loan, which meets the ListingCategory. So how about the borrowers' income:
```{r echo=FALSE, Univariate_Plots}
prosperLoan$IncomeRange = factor(prosperLoan$IncomeRange, 
                                 levels=c("Not employed", "$0", "$1-24,999", "$25,000-49,999", 
                                          "$50,000-74,999", "$75,000-99,999", "$100,000+", "Not displayed"))

ggplot(data = prosperLoan, aes(x = IncomeRange)) + 
  geom_bar(color = "black", fill = "pink") + 
  theme(axis.text.x  = element_text(angle=30, vjust=0.5)) +
  ggtitle("Income range of the borrowers")
```
Most of the people using Prosper have an income between $25000 and $75000. The original amount of a loan can definitely impact the borrower rate, so let's study it.
```{r echo=FALSE, Univariate_Plots}
qplot(x = LoanOriginalAmount, data = prosperLoan, bins = 30)
```
It looks like a long tail data, so for better understand the distribution of the LoanOriginalAmount, I transformed this distribution on a log scale.
```{r echo=FALSE, Univariate_Plots}
qplot(x = LoanOriginalAmount, data = prosperLoan, bins = 1000) + 
  scale_x_log10()
```
We see amounts are 1000, 2000, 5000, and 10000, these numbers are the most. I also created a new variable TotalProsperPaymentdelinquencies which is equal to the difference between TotalProsperPaymentsBilled and OnTimeProsperPayments. If a borrower trend to delay his payments of loan, I think Prosper will give him an higher borrower rate when he wanna borrow a new loan next time. But I realized it's quite difficult to discuss it, because I need to calculate the borrowers who has borrowed a loan from Prosper for at least 2 times, then analysis the influence between the latter borrowerRate and the former paymentdelinquencies. Actually this variable TotalProsperPaymentdelinquencies is more useful for judging how easliy one person can borrow a loan form Prosper. So how about the currentDelinquencues:
```{r echo=FALSE, Univariate_Plots}
qplot(x = CurrentDelinquencies, data = prosperLoan) + 
  scale_x_continuous(limits = c(-1, 33),breaks = seq(-1,33,3)) +
  scale_y_continuous(limits = c(-1, 12000),breaks = seq(-1,12000,1000))
```
A large part of people haven't had a delinquencies, while there are serval borrowers have had one or two times delinqunencies, also very few people had more than 10 times delinquencies, I'm curious about how can they get further loans with such a bad credit, maybe their further loans should provide higher borrower rate.

# Univariate Analysis

### What is the structure of your dataset?
There are 113937 loan records with 81 variables in the dataset, some are numeric and some are factor. While I choose 15 variables which are all relative to the borrowers from the whole set.

### What is/are the main feature(s) of interest in your dataset?
I care about what kinds of people will borrow a loan and what characters can improve the deal of loan. So the main feature in this data set is borrowerrate, which a borrower cares most. I want to find out which variables have impact on it.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
I think that ProsperRating and CreditGrade have big impacts on BorrowerRate, since lending money to lower rating borrower may have higher risks. Moreover, IncomeRange, Term, Occapution, CurrentDelinquencies and LoanOriginalAmount may also help me investigate the BorrowerRate.

### Did you create any new variables from existing variables in the dataset?
Yes, The variable 'TotalProsperPaymentdelinquencies' represents the delinquencies a borrower made, which is equal to the difference between 'TotalProsperPaymentsBilled' and 'OnTimeProsperPayments', because I think this variable may represent a borrower's credit, which is very important for applying a loan. Also I transform the value of 'ListingCategory' into the category which it means. And transform the value of 'ListingCreationDate' into date value.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
Yes, I divied the variable ListingCreationDate into year and month. When I drawed the ListingCreationDate into histogram, I found in each year there is a trend growing higher, so I guess that maybe in different months, the desire of people wanting to borrow a loan is changeable.


# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}
ggplot(data = prosperLoan, aes(x = ProsperScore, y = BorrowerRate)) +
        geom_boxplot(aes(fill = ProsperScore)) +
        xlab("Score") +
        ggtitle("Yield for Different Prosper Score")

qplot(x = ProsperScore, y = BorrowerRate,
      data = subset(prosperLoan, !is.na(ProsperScore)), 
      geom = 'boxplot')

by(prosperLoan$BorrowerRate, prosperLoan$ProsperScore, summary)

```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?




# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
