---
title: "Prosper Loan Exploration"
author: "Samuel Wang"
date: "2017-01-05"
output: html_document
---

How easily you can borrow loan from Prosper
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.
install.packages("ggplot2", dependencies = T) 
install.packages("knitr", dependencies = T)
install.packages("dplyr", dependencies = T)


library(ggplot2)
library(plyr)
```
This report explores a dataset containing LoanStatus and attributes for almost 114000 loan data in Prosper. I pick out 15 variables from the total 81 variables which I think are strongly relative to borrower. So firstly I deal with the original data into a subset and take a look at the subset as follows:
```{r echo=FALSE, Load_the_Data}
# Load the Data
prosperLoan <- read.csv(file = "C:\\Users\\wangs\\Downloads\\subsetProsperLoanData.csv", header = TRUE, sep = ",")

subvariables <- c('ListingCreationDate','CreditGrade','Term','LoanStatus','BorrowerRate','ProsperRating..Alpha.','ProsperScore','ListingCategory..numeric.','BorrowerState','Occupation','CurrentDelinquencies','IncomeRange','TotalProsperLoans','TotalProsperPaymentsBilled','OnTimeProsperPayments','LoanOriginalAmount')

prosperLoan <- subset(prosperLoan,select=subvariables)

colnames(prosperLoan)[6] <- "ProsperRating"
colnames(prosperLoan)[8] <- "ListingCategory"

prosperLoan$TotalProsperPaymentdelinquencies <- prosperLoan$TotalProsperPaymentsBilled - prosperLoan$OnTimeProsperPayments

subvariables <- c('ListingCreationDate','CreditGrade','Term','LoanStatus','BorrowerRate','ProsperRating','ProsperScore','ListingCategory','BorrowerState','Occupation','CurrentDelinquencies','IncomeRange','TotalProsperLoans','TotalProsperPaymentdelinquencies','LoanOriginalAmount')

prosperLoan <- subset(prosperLoan,select=subvariables)

prosperLoan[is.na(prosperLoan)] <- 0

ListingCategoryMap <- c("0"="Not Available","1"="Debt Consolidation","2"="Home Improvement","3"="Business",
                        "4"="Personal Loan","5"="Student Use","6"="Auto","7"="Other","8"="Baby&Adoption",
                        "9"="Boat","10"="Cosmetic Procedure","11"="Engagement Ring","12"="Green Loans",
                        "13"="Household Expenses","14"="Large Purchases","15"="Medical/Dental","16"="Motorcycle",
                        "17"="RV","18"="Taxes","19"="Vacation","20"="Wedding Loans")

prosperLoan$ListingCategory <- as.character(prosperLoan$ListingCategory)
revalue(prosperLoan$ListingCategory,ListingCategoryMap)

prosperLoan$ListingCategory[prosperLoan$ListingCategory == "0"] <- "Not Available"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "1"] <- "Debt Consolidation"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "2"] <- "Home Improvement"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "3"] <- "Business"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "4"] <- "Personal Loan"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "5"] <- "Student Use"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "6"] <- "Auto"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "7"] <- "Other"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "8"] <- "Baby&Adoption"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "9"] <- "Boat"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "10"] <- "Cosmetic Procedure"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "11"] <- "Engagement Ring"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "12"] <- "Green Loans"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "13"] <- "Household Expenses"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "14"] <- "Large Purchases"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "15"] <- "Medical/Dental"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "16"] <- "Motorcycle"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "17"] <- "RV"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "18"] <- "Taxes"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "19"] <- "Vacation"
prosperLoan$ListingCategory[prosperLoan$ListingCategory == "20"] <- "Wedding Loans"

prosperLoan$ListingCategory <- as.factor(prosperLoan$ListingCategory)

prosperLoan$ListingCreationDate <- as.Date(prosperLoan$ListingCreationDate, format='%Y-%m-%d')

write.csv(file = './subsetProsperLoanData.csv', prosperLoan, na="0")

names(prosperLoan)
```
The variable 'TotalProsperPaymentdelinquencies' represents the delinquencies a borrower made, which is equal to the difference between 'TotalProsperPaymentsBilled' and 'OnTimeProsperPayments', because I think this variable may represent a borrower's credit, which is very important for applying a loan. Also I transform the value of 'ListingCategory' into the category which it means. And transform the value of 'ListingCreationDate' into date value.

Now let's explore it. I'd like to see the user counts at different years. We could see that the loan records are distributed from the year 2006 to 2014, while there is a gap at 2009, I guess that maybe there is an update of the system and the data before 2009 is on the old one. Also I see that the records are much more in 2014 than the former years. 
```{r echo=FALSE, Load_the_Data}
qplot(x = ListingCreationDate, data = prosperLoan, bins = 120)
```
Maybe there is a relationship between the month and the loan record, so I divide the variable 'ListingCreationDate' into year and month to see what will show, well, it seems there is no clear evidence.
```{r echo=FALSE, Load_the_Data}
prosperLoan$year <- as.numeric(format(prosperLoan$ListingCreationDate, '%Y'))
prosperLoan$month <- as.numeric(format(prosperLoan$ListingCreationDate, '%m'))

qplot(x = month, data = prosperLoan) +
  scale_x_continuous(breaks = seq(0,12,3)) +
  facet_wrap(~prosperLoan$year, scales = "free_y", ncol = 5)
```
The length of the loan expressed in months. And we can find that Prosper provides mostly 36-months loans.
```{r echo=FALSE, Load_the_Data}
prosperLoan$Term<-factor(prosperLoan$Term)
ggplot(data=prosperLoan,aes(Term)) + 
  geom_bar(color=I('black'),fill=I('blue'))
```
We can see that most of the loans are active but there still are some defaluted payments.
```{r echo=FALSE, Load_the_Data}
p <- ggplot(prosperLoan, aes(LoanStatus))
p <- p + geom_bar(width = 1, colour="black", fill="blue")
p <- p + geom_text(aes(y = ..count.., label = ..count..),
                   stat = "count", color = "white",
                   hjust = 1.0, size = 3)
p <- p + theme(legend.position = "none")
p <- p + coord_flip()
p
```
The Borrower Rate seems quite uniformly distributed.
```{r echo=FALSE, Load_the_Data}
ggplot(data=prosperLoan,aes(BorrowerRate)) + 
  geom_histogram(color=I('black'),fill=I('#56B4E9'))
```

```{r echo=FALSE, Load_the_Data}
prosperLoan$IncomeRange = factor(prosperLoan$IncomeRange, 
                                 levels=c("Not employed", "$0", "$1-24,999", "$25,000-49,999", 
                                          "$50,000-74,999", "$75,000-99,999", "$100,000+", "Not displayed"))

ggplot(data = prosperLoan, aes(x = IncomeRange)) + 
  geom_bar(color = "black", fill = "blue") + 
  theme(axis.text.x  = element_text(angle=30, vjust=0.5)) +
  ggtitle("Income range of the borrowers")
```

Our dataset consists of 15 variables, with around 114,000 obseravations.

# Univariate Plots Section
```{r echo=FALSE, Univariate_Plots}
ggplot(aes(x = ProsperScore), data = prosperLoan) + 
  geom_histogram()

p <- ggplot(prosperLoan, aes(LoanStatus))
p <- p + geom_bar(width = 1, colour="white")
p <- p + geom_text(aes(y = ..count.., label = ..count..),
                   stat = "count", color = "white",
                   hjust = 1.0, size = 3)
p <- p + theme(legend.position = "none")
p <- p + coord_flip()
p

qplot(x = ListingCreationDate, data = prosperLoan)

qplot(x = CreditGrade, data = prosperLoan)

qplot(x = Term, data = prosperLoan)

qplot(x = BorrowerRate, data = prosperLoan, binwidth = 0.001)

qplot(x = ProsperRating, data = prosperLoan)

qplot(x = ProsperScore, data = prosperLoan)

prosperLoan <- within(prosperLoan,
                      ListingCategory <- factor(ListingCategory, 
                                                levels = names(sort(table(ListingCategory),decreasing = TRUE))))
p <- ggplot(prosperLoan, aes(ListingCategory))
p <- p + geom_bar(width = 1, colour="white")
p <- p + geom_text(aes(y = ..count.., label = ..count..),
                   stat = "count", color = "white",
                   hjust = 1.0, size = 3)
p <- p + theme(legend.position = "none")
p <- p + coord_flip()
p


prosperLoan <- within(prosperLoan,
                      BorrowerState <- factor(BorrowerState, 
                                                levels = names(sort(table(BorrowerState),decreasing = TRUE))))
p <- ggplot(prosperLoan, aes(BorrowerState))
p <- p + geom_bar(width = 1, colour="white")
p <- p + geom_text(aes(y = ..count.., label = ..count..),
                   stat = "count", color = "white",
                   hjust = 1.0, size = 3)
p <- p + theme(legend.position = "none")
p <- p + coord_flip()
p


prosperLoan <- within(prosperLoan,
                      Occupation <- factor(Occupation, 
                                                levels = names(sort(table(Occupation),decreasing = TRUE))))
p <- ggplot(prosperLoan, aes(Occupation))
p <- p + geom_bar(width = 1, colour="white")
p <- p + geom_text(aes(y = ..count.., label = ..count..),
                   stat = "count", color = "white",
                   hjust = 1.0, size = 3)
p <- p + theme(legend.position = "none")
p <- p + coord_flip()
p


qplot(x = CurrentDelinquencies, data = prosperLoan)

qplot(x = IncomeRange, data = prosperLoan)

qplot(x = LoanOriginalAmount, data = prosperLoan)

qplot(x = LoanOriginalAmount, data = prosperLoan) + 
  scale_x_continuous() +
  scale_x_log10()

qplot(x = LoanOriginalAmount, data = prosperLoan)
```

# Univariate Analysis

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?



# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?




# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
